---
title: "10-import-data"
output:
  html_document: 
    code_folding: show
    theme: cosmo
    toc: yes
    toc_depth: 5
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r load libraries }
# Use pacman, which forces an install if the library isn't present on the running machine
if (!require("pacman"))
  install.packages("pacman")

# Load libraries required for this file
pacman::p_load(tidyverse, janitor, haven, furrr, readxl)
```

# Importing data
The purpose of this file is to contain all of the functions required or used to generate the dataset which will be investigated.  What follows are functions to import and generate data, and also pre-constructed lists for a simplified selection of features from the dataset.

## Sample workflow
The following code chunk defines the expected way that the following functions will be used, although other workflows are possible:

```{r, example chunk, eval=FALSE, purl=FALSE}
## load waves and join them
wave_data <- load_waves(1:5)
full_dataset <- get_working_dataset_full(wave_data, join_type = 'full')

## set outcome variable of interest
outcome = 'h5mn8'

## get the aids that you want
inner_aids <- get_inner(list(wave_data[[1]], wave_data[[3]], wave_data[[4]], wave_data[[5]]))

## use the features and ids that you want to select out what you want
working_ds <- full_dataset %>%
  filter(aid %in% inner_aids) %>%
  dplyr::select(aid, predictor_list, outcome)
```



## Public helper functions
These functions are helpers in case you need to know something about your dataframe more than what is provided in the general loading/joining interface.  Use as desired!

```{r generate unique list of vars from df}
# Return all unique values of var_name in data frames
get_unique_varnames <- function(df_list, var_name='aid'){
  # df_list : list of dataframes
  # var_name: variable of interest.  Expected in all lowercase letters
  # returns: list of distinct values of variable of interest across list of dataframes
  if(length(df_list)==0)
    stop("You didn't pass in any dataframes!")
  
  # Get list of var_names
  var_list <- df_list %>%
    bind_rows() %>%
    dplyr::select(var_name)
  
  # Return the unique ones
  return (distinct(var_list))
}
```

```{r check variable existence}
# Function to Check Existence of Variable in Data
# Use this function to check if a certain variable exists in a given. 
var_in_df <- function(df, varname){
  # df : dataframe to check in
  # varname : varaible to check for
  # return : Bool indicatind if the variable exists in the dataframe or not
  any(colnames(df) == varname)
}
## Examples of Usage : joined_waves %>% var_in("h4id5j")
```

```{r count responses by variable of interest}
# Function to check the counts by categories in the variables
count_response_by_variable <- function(df, ...){
  # df : dataframe where the variable exists
  # ... : variable you want to check the counts of - sent in as a string
  df %>%
    group_by_(...) %>% 
    summarise(n = n())
}

# Example of usage
# Tibble for first outcome with all of its predictors
# suicide_df <- joined_waves %>%
#   get_outcomes_df(outcome = "h5mn8", 
#                   predictors = setdiff(predictor_list, "aid"))
```

```{r returns intersection of variables in dataframes list}
# This function returns values that are common to all dataframes in list; 'aid' might be the only one here.
# Errors out if the variable isn't in all frames.

get_inner <- function(in_list, var_name='aid'){
  #in_list : contains lists of tibbles of wave data
  #var_name: column name of interest
  #returns: list of elements in the var_name column that are common to all of the dataframes
  all_var_list <- in_list %>%
    map(dplyr::select, var_name) %>%
    purrr::reduce(intersect) %>%
    pull(var_name)
  
  return(all_var_list)
}

## Example of usage
# wave_data <- load_waves(1:5)
# inner_aids <- get_inner(list(wave_data[[1]], wave_data[[3]], wave_data[[4]], wave_data[[5]]))
```

## Private helper functions
These helper functions are used within the publicly available functions.  Usage of these functions directly should be avoided.

## Public functions for loading data
These functions provide the overall functionality for loading a single file, multiple wave files, or multiple files in general using the dialog chooser.  Keep in mind that you can only use the dialog chooser on windows.  Use all functions as desired!

```{r load single file}
# This helper loads a single file given a full filepath f_path;

load_file <- function(f_path, use_clean_names = TRUE) {
  # f_path : filepath of AddHealth xpt file
  # returns: tibble of read data
  
  # Read the data
  read_data <- read_xpt(f_path)
  
  # Figure out what kind of data it is (currently race or a wave file)
  f_name = basename(f_path)
  
  # If it's wave data, extract the wave number and append it to the end of the dataframe.  Otherwise, read as normal.
  # We need to think through the consequences of this action.
  # if (str_extract(f_name, 'wave') == 'wave') {
  #   
  #   wave_no = parse_number(f_name)
  #   
  #   read_data <- read_data %>%
  #     mutate(wave_no = wave_no)
  # }
  
  # If you want clean names
  if(use_clean_names){
    read_data <- read_data %>%
      clean_names()
  }
  
  return(read_data)
}
```

```{r load using file chooser}
# This function allows choosing of files from a file chooser; returns a list containing the files.  Only available for Windows.

load_using_chooser <- function(use_clean_names = TRUE) {
  # use_clean_names: boolean of whether to fix column names
  # returns: list of tibbles of read data
  
  # Generate useful error for users not on Windows
  if(.Platform$OS.type != 'windows')
    stop(str_c("Sorry, this functionality isn't available on ", .Platform$OS.type))
  
  # Generate file chooser so wave or other data can be selected via GUI
  files <- choose.files(
    default = "",
    caption = "Select files",
    multi = TRUE,
    filters = Filters,
    index = nrow(Filters)
  )
  
  # Load data; returned as list of tibbles
  file_data <- files %>%
    map(load_file, use_clean_names)
  
  return(file_data)
}
```

```{r load all wave files easily}
# This function allows loading of files from a set location based on the waves of interest.

load_waves <- function(wave_no_list, use_clean_names = TRUE, filebase='~/gvfs/smb-share:server=vu1file.it.vanderbilt.edu,share=gaydosh_addhealth'){
  # wave_no_list : list of waves you want to load
  # use_clean_names: boolean of whether to fix (e.g., lowercase) column names
  # filebase : custom filepath for your local filesystem if needed; directly precedes .../Gaydosh/Core Files/...
  # returns: list of tibbles of wave files
  
  # Default file ending
  file_end = '/Gaydosh/Core Files/In Home Interview Files/wave'
  
  #Create filepath
  filepath = str_c(filebase, file_end)
  
  # Generate wave file list
  files <- str_c(filepath, wave_no_list, ".xpt")
  
  # Load data; returned as lists
  file_data <- files %>%
    future_map(load_file, use_clean_names)
  
  return(file_data)
}

# Example usage
# wave_data <- load_waves(1:5)
```

## Public functions for joining data and selection of data
These functions allow for joining data using different types of joins, and also allow for subsets of the data to be selected based on the input features
TODO:  We may need to think about what happens with the suffix - how will repeated variables (although we don't have that problem) be renamed?
TODO:  If you have names that are repeated, this will not work correctly because the variables will have repeated names (e.g., `h1fs1.x`) which won't match `h1fs1`.
TODO:  `get_working_dataset` requires assistance in determining how we can figure out which variables are factors without the codebook; note that somehow the `mutate_at` statement doesn't work in its current formulation (using | bars instead of commas) on ACCRE...

```{r perform join of data in an expedited way}
#This function returns the dataset with the selected predictors based on your join strategy.  However, with this function, you cannot easily retrieve the rows you left out without another join.

get_working_dataset_exp <- function(df_list, feature_list = NULL, join_type = 'inner', by='aid'){
  # df_list : list of dataframes
  # feature_list : list of desired features.  If null, select all features.
  # join_type : type of join you'd like to perform
  # by: key on which the join should be performed
  # returns: full joined tibble with selected features
  
  # Obtain correct join type based on inputs
  if(join_type=='inner')
    jtype <- inner_join
  else if(join_type=='left')
    jtype <- left_join
  else if(join_type=='right')
    jytpe <- right_join
  else if(join_type=='full')
    jtype <- full_join
  else{
    stop(str_c('Join type', by, 'is not supported.', sep=' '))
  }
  
  # Do the join
  ds <- reduce(df_list, jtype, by=by)
  
  # Remedy null feature list if null
  if (is.null(feature_list)){
    feature_list = colnames(ds)
  }
  
  # Make sure all values in feature list are in the column names
  missing_features <- setdiff(feature_list, colnames(ds))
  
  if(length(missing_features) != 0){
    stop('The following features are missing in the set of dataframes you passed in: ', str_c(missing_features, collapse=', '))
  }
  
  # Fix factor issues and dplyr::select features
  ds <- ds %>% 
    mutate_at(vars(starts_with("h"), starts_with("r"), starts_with("m")), as_factor) %>%
    dplyr::select(all_of(feature_list))
  
  return(ds)
}

```

This function has the same problems as `get_working_dataset_exp`.  However, this function may end up being more useful depending on the file format that we decide to save the fully joined data in (if we can).  The reading of the remote file is still TODO.
TODO: Something is strange with my mutate_at function; it can only work with commas and not bars.

```{r perform join of data}
# This functions generates the data by starting with an outright full join of the data

get_working_dataset_full <- function(df_list, feature_list = NULL, join_type = 'inner', load_join=FALSE){
  # df_list : list of dataframes
  # feature_list : list of desired features.  If null, select all features.
  # join_type : type of join you'd like to perform
  # by: UNIMPLEMENTED, TODO
  # load_join: if true, the fully joined data will be loaded (UNIMPLEMENTED, TODO)
  # returns: full joined tibble with selected features
  
  # Do the full join
  if(load_join){
    # Here, we want to load the joined data from file if possible and contractually acceptable.
  }
  else
    ds <- reduce(df_list, full_join, by='aid')
  
  # Extract AIDs from each df
  aid_lists <- df_list %>% map(dplyr::select, aid)
  
  # Obtain correct join type based on inputs
  if(join_type=='inner')
    aids <- reduce(aid_lists, intersect) %>% pull(aid)
  else if(join_type=='left')
    aids <- aid_lists[[1]] %>% pull(aid)
  else if(join_type=='right')
    aids <- aid_lists[[length(aid_lists)]] %>% pull(aid)
  else if(join_type=='full')
    aids <- reduce(aid_lists, union) %>% pull(aid)
  else{
    stop(str_c('Join type', by, 'is not supported.', sep=' '))
  }
  
  # Filter out aids
  ds <- ds %>%
    filter(aid %in% aids)
  
  # Remedy null feature list if null
  if (is.null(feature_list)){
    feature_list = colnames(ds)
  }
  
  # Make sure all values in feature list are in the column names
  missing_features <- setdiff(feature_list, colnames(ds))
  
  if(length(missing_features) != 0){
    stop('The following features are missing in the set of dataframes you passed in: ', str_c(missing_features, collapse=', '))
  }
  
  # Fix factor issues and select features
  ds <- ds %>% 
    #mutate_at(vars(starts_with("h"), starts_with("r"), starts_with("m")), as_factor) %>%
    dplyr::select(all_of(feature_list))
  
  return(ds)
}

# Example usage:
# wave_data = load_data(1:5)
# full_dataset <- get_working_dataset_full(wave_data, join_type = 'full')
```

TODO: We need to make some decisions about how demographic data will likely be included in the generated data.  Where are these variables included in the variable lists and how can we most intuitively integrate them into our data loading?
```{r}

# This function will be used to add age and race demographic into the data
add_demographics <- function(df){
  # df : joined waves input
  # return : new data frame with race and age information at each wave. 
  #          age_w{x} - where x is the wave number (e.g. age_w5, age_w4..etc)
  # race info comes from wave 5
  # if h1giy == 96, then age is marked as NA ~ 17 people refused providing their ages
  ## Details:
  ## h1gi1y is the birth year of the person 
  ## iyear3,iyear4,iyear5 are the years when the inerviews are conducted in waves 3, 4 and 5
  ## details on age calculation provided in comments below.
  # return : df with age and race demographic added.
  
  df %>% 
    dplyr::mutate(
      
      # We calculate the birth of the individual in months from year 0 as a reference. 
      # h1gi1y is a two digit representation of the year of birth 
      ## - if a person is born in 87 this gets concatenated with 19 to make 1987 and then converted to a double to calculate the birthdate in months. 
      ## - if the value of h1gi1y is given as 96, then it is marked as NA and the age of that person will be carried over as NA across all the other waves.
      
      birth_in_months = ifelse(h1gi1y != 96, (as.double(str_c("19",h1gi1y))*12 )+ as.double(h1gi1m), NA), 
      
      # AGE CALCULATION DETAILS: This is the logic used to calculate age_w1, age_w2, age_w3, age_w4, age_w5
      #
      ## age_w1 is the age of the individual at wave 1.
      ##
      ## This is calculated by:
      ### 1. finding the interview year (be careful with this as this could be a two digit representation of the year)
      #### - if this is in string format, it will need to be concated named with 19 and then changed to double for calculation
      ### 2. Multiplying that interview year by 12 to get the total number of months since year 0
      ### 3. Adding the month of the interview to this to get the interview date IN MONTHS
      ### 4. Subtract the birthdate in months and divide by 12 to get the age of individual at time of interview
      ### 5. round to 1 decimal point so if the person is aged 37.89, their age is read in as 37.9. 
      
      age_w1 = round((((as.double(str_c("19",iyear))*12 )+imonth) - birth_in_months)/12, 1), # iyear was given as a two digit year here, hence it was concatenated with 19
      
      age_w3 = round((((iyear3*12)+imonth3)-birth_in_months)/12, 1),
      
      age_w4 = round((((iyear4*12)+imonth4)-birth_in_months)/12, 1),
      
      age_w5 = round((((iyear5*12)+imonth5)-birth_in_months)/12, 1)) %>% 
    
    dplyr::select(-one_of(c('birth_in_months'))) %>%
    dplyr::mutate(one_race5 = case_when( #one race: first analyze those with multiple races
      (h5od8==1) ~ 1, #1 = white
      (h5od8==2) ~ 2, #2 = black
      (h5od8>=3)&&(h5od8<=7) ~ 3, #3= hispanic
      (h5od8>=9)&&(h5od8<=15) ~ 4, #4 = asian
      (h5od8>=16)&&(h5od8<=22) ~ 5, #5 = other
      #then analyze only 1 race response
      (h5od4a==1) ~ 1, #1 = white
      (h5od4b==1) ~ 2, #2 = black
      (h5od4c==1) ~ 3, #3= hispanic
      (h5od4d==1) ~ 4, #4 = asian
      ((h5od4e==1)|(h5od4f==1)|(h5od4g==1)) ~ 5)) #other
  #otherwise NA default
}

```


The following function is used because for some reason, there are some subjects in Waves 2 and 3 that are not in Wave 1.  The following function loads the errant AIDs from a remote server and then removes them from the desired dataset.  Details of this function in file `project-component-verification.Rmd`
```{r remove errant subjects}
# This function removes subjects that are not present in Wave 1
remove_subjects_not_in_wave1 <- function(df, filebase = '~/gvfs/smb-share:server=vu1file.it.vanderbilt.edu,share=gaydosh_addhealth'){
  #df: dataset of interest with AIDs as column
  #returns: df with errant subjects removed
  
  # Default file ending
  file_end = '/DSI/exclusion_aids.csv'
  
  #Create filepath
  excl_fname = str_c(filebase, file_end)
  
  #Get the total number of subjects for usage later
  n_subjs <- nrow(df)
  
  #Get bad aids fom file
  excl_aids <- read_csv(excl_fname, col_types=cols(col_character()))
  
  #Remove bad aids from dataset
  filtered_ds <- df %>%
    filter( !(aid %in% pull(excl_aids)) )
  
  #Tell the user about the results of this function
  warning(str_c(n_subjs - nrow(filtered_ds), " subjects removed from dataset.\n"))
  
  return(filtered_ds)
}
```


## Recode Miss Variables
Levels of question that were "miss" variables. For numeric they are coded as NA and then later dropped, for factors they are coded as just coded as NA.
```{r}
recode_missing_factors<-function(df, na_levels){
  
  print("Recoding Missing Factor Variables")
  # function to replace skip levels in factor variables with NA
  # df: dataframe of wave data
  # na_levels: read data from na_levels.csv file
  
  
  
  var_names <- intersect(na_levels %>%  
                           filter(type == "Factor") %>%
                           select(variable_name) %>% 
                           pull() %>% tolower(), names(df))
  
  print(paste("Number of factor variables being recoded : ", length(var_names) ))
  print("Factor variables being recoded : ")
  print(var_names)
  
  for (i in 1:length(var_names)){
    
    print(var_names[i])
    mlevels <- na_levels %>% 
      filter(variable_name == var_names[i]) %>% 
      select(na_values) %>% 
      pull() %>% 
      str_split(',') %>% 
      .[[1]] %>% 
      as.character()
    
    if(!is.na(mlevels[1])){ 
      
      if (class(df[[var_names[[i]]]]) != "factor"){ # if the variable to be changed is not a factor already
        df[[var_names[[i]]]] <- df[[var_names[[i]]]] %>% as_factor() 
      }
      
      levels(df[[var_names[i]]])[levels(df[[var_names[i]]]) %in% mlevels] <- NA
    }

  }
  
  df[var_names] <- lapply(df[var_names] , factor)
  print(sapply(df[var_names], class))
  
  return(df)
}

#Example
 #csv containing NA descriptions
# na_csv <- read_csv('na_levels.csv')
# #Test tibble
# test_df <- tibble(
#   h1fs1 = c(1,1,2,2,3,4,5,6,7,8,9,9,1,1),
#   h1fs3 = c(2,3,1,2,3,4,6,3,8,2,9,9,1,NA),
#   h3wgt = c(2,888, 4, 6, 2, NA, 8, 9, 12, 24, 11, 12, 996, 33)
# )
# recode_missing_factors(test_df, na_csv)
```


```{r}
recode_missing_numeric<-function(df, na_levels){
  # function to replace skip levels in factor variables with NA
  # df: dataframe of wave data
  # na_levels: read data from na_levels.csv file
  
  print("Recoding Missing Numeric Variables")
  var_names <- intersect(na_levels %>%
                           filter(type == "Numeric") %>% 
                           select(variable_name) %>% 
                           pull() %>% 
                           tolower(), names(df))
  
  print(paste("Number of numeric variables being recoded : ", length(var_names) ))
  print("Numeric variables being recoded : ")
  print(var_names)
  
  
  for (i in 1:length(var_names)){

    mlevels <- na_levels %>% 
      filter(variable_name == var_names[i]) %>% 
      select(na_values) %>% 
      pull() %>% 
      str_split(',') %>% 
      .[[1]]  

    
    if(!is.na(mlevels[1])){ # verify that the variable in question actually has missing levels
      
      df[,var_names[i]] <- ifelse(as.vector(pull(df[,var_names[i]]) %in% mlevels),
                                       NA, 
                                       pull(df[,var_names[i]]))
      # if the column of the numeric variable has values that are supposed to be skipped
      # replace them with NA
      # else keep the original value
      
    }
    
  }
  
  
  return(df)
}

# Example
# # read csv file of NA values
# na_csv <- read_csv('na_levels.csv')
# 
# #Test tibble
# test_df <- tibble(
#   h1fs1 = c(1,1,2,2,3,4,5,6,7,8,9,9,1,1),
#   h1fs3 = c(2,3,1,2,3,4,6,3,8,2,9,9,1,NA),
#   h3wgt = c(2,888, 4, 6, 2, NA, 8, 9, 12, 24, 11, 12, 996, 33)
# )
# 
# recode_missing_numeric(test_df, na_csv)

```


```{r}
recode_variables <- function(df, sheet_num, outcome = NULL, binarize = TRUE){
  # Function to recode variables from different sheets in the variables_of_interest file
  ## we will use this function to load different variabless: despair predictors, relevant individual predictors, contextual predictors, outcome variables
  # df: dataframe of wave data
  # sheet_num : this sheet number
  # outcome : outcome variable of interest that needs to be recoded
  # binarize : if TRUE means that the outcome variable has a range to be binarized to 0 or 1
  
  
  sheet_df <- read_excel("Variables_of_Interest.xlsx", sheet = sheet_num) %>% clean_names()
  sheet_df$variable_name <- sheet_df$variable_name %>% tolower()
  
  #if you're recoding the predictors
  if(is.null(outcome)){
    
      df <- recode_missing_factors(df, sheet_df) %>%
        recode_missing_numeric(sheet_df)
  
  #otherwise you're recoding the outcomes    
  } else {
    
    # if they're not constructed variables, recode them according to sheet
    if ((outcome %in% constructed_outcomes) == FALSE) {
      
      # dropping rows where outcoome variable is NULL
      df <- recode_missing_factors(df, sheet_df) %>% drop_na(outcome) 
      
      #if the values have a range and you need to binarize them to 0 or 1:
      if (binarize == TRUE) {
        fct_levels <- df %>%
          select(outcome) %>%
          pull() %>%
          unique() %>%
          as.vector()
        
        not_zero <- fct_levels[fct_levels != 0]
        
        df[outcome] <-
          fct_collapse(df[outcome] %>% pull() %>% as.vector(), "1" = not_zero)
        
        print(nrow(df))
      }
      
    } else {

      # convert constructed outcome to factors otherwise and drop NA
      df <- df %>%
        mutate(!!outcome := as.factor(.data[[outcome]])) %>%
        drop_na(outcome)
    }
    
  }
  
  return(df)
}
```


## Lists of Selected Predictor Variables
The following are some lists of predictor variables, organized by area of interest.  Note that the AIDs have been removed from all of these.

```{r lists of predictors for easy study}
# List of all emotional despair measures

diagnosed_depression_list <- c("h4id5h","h5id6g")

depression_list <- c(
  "h1fs1","h1fs3","h1fs4","h1fs5","h1fs6","h1fs7","h1fs11","h1fs15","h1fs16","h1fs17", 'h1fs19',   # Wave 1 ~ 10 vars
  "h3sp5","h3sp6","h3sp7","h3sp8","h3sp9","h3sp10","h3sp11","h3sp12","h3sp13",           # Wave 3 ~ 9 vars
  "h4mh18","h4mh19",'h4mh20','h4mh21',"h4mh22","h4mh23","h4mh24","h4mh25", "h4mh26","h4mh27",      # Wave 4 ~ 9 vars
  "h5ss0a","h5ss0b","h5ss0c","h5ss0d", "h5ss0e", "h5id6i")                                # Wave 5 ~ 6 vars

# List of all Anxiety Variables - Note only wave 4 is being looked at for Anxiety measures
diagnosed_anxiety_list <- c("h4id5j","h5id6i")
anxiety_list <- c("h4pe6","h4pe14","h4pe22","h4pe30")

emotional_despair_predictors <- c(diagnosed_depression_list, depression_list, diagnosed_anxiety_list, anxiety_list)

# List of all cognitive despair measures
optimism_list <- c("h4pe7","h4pe15","h4pe23","h4pe31","h5pe1","h5pe2", "h5pe3")
other_cog_despair_list <- c('h4mh3', 'h4mh4', 'h4mh5', 'h4mh6', 'h4mh2',
                            'h5mn1', 'h5mn2', 'h5mn3', 'h5mn4')
future_hopes_list <- c("h1ee14", "h1ee12", "h3ec56")

cognitive_despair_predictors <- c(optimism_list, other_cog_despair_list, future_hopes_list)

#List of all biological despair measures

bio_despair_predictors <- c(#'h3wgt', 'h3hgt_f', 'h3hgt_i', these will be used in calculations for h3bmi 
'h3bmi', # this will be the contructed variable (wave 3 bmi) from the above three variables
'h4bmi', 'hba1c', 'crp', 'hdl', 'ldl', 'tg', 'h4waist', 'h4sbp', 'h4dbp', 'h4bpcls',
'h5bmi', 'h5waist', 'h5sbp', 'h5dbp', 'h5bpcls4', 'h5bpjcls')


bio_despair_core_list_exp <- c('h1gh60', 'h1gh59a', 'h1gh59b', 'h3wgt', 'h3hgt_f', 'h3hgt_i', 'h4bmi',
                               'h4waist', 'h4sbp', 'h4dbp', 'h4bpcls', 'h5bmi', 'h5waist', 'h5sbp', 
                               'h5dbp', 'h5bpcls4', 'h5bpjcls')

bio_despair_other_list_exp <- c('hba1c', 'crp', 'hdl', 'ldl', 'tg')

# List of all behavioral despair measures
behavioral_despair_predictors <- c("h1ds14", "h1ed9", "h1ed7", "h1ds3", "h1ed2", "h1fv7", "h1fv1", "h1fv8", "h1jo9", "h1ds13", "h1ds12", "h1ds11", "h3ds18h", "h3ds18a", "h3ds18i", "h3to49", "h3ds6", "h3ds5", "h3ds4", "h4ds7", "h4ds19", "h4ds14", "h4ds20", "h4ds6", "h4ds5", "h4ds4", "h4cj17", "h5cj1d", "h5cj1e", "h5cj2b", "h5cj1f", "h5cj1b", "h5cj1c", "h5cj1a")

# List of all Demographic Variable
demographic_list <- c("one_race5","bio_sex") #removed "race1","mulrace1","race1rul"

demographic_age_list <- c("age_w5") #"age_w1", "age_w3", "age_w4" ## Removing ages 1-4. 

# Total predictor variable list
predictor_list <- c(emotional_despair_predictors, cognitive_despair_predictors, bio_despair_predictors, behavioral_despair_predictors)

# Note: Look into importing these later
#c4var009,
#c4var010,
#c4var002
```

###Predictor list addendum
Currently, we're not sure about what files some of the variables are located in.  For now, we're just going to remove those errant variables from the predictor list
```{r unsure variables}
# unsure_variables <- c('hba1c', 'crp', 'hdl', 'ldl', 'tg', 'h5bmi', 'h5waist', 'h5sbp', 'h5dbp', 'h5bpcls4', 'h5bpjcls')
# predictor_list <- setdiff(predictor_list, unsure_variables)
```

# Lists of Selected Outcome Variables
The following are some lists of outcome variables, organized by area of interest.  Note that the AIDs have been removed from all of these.
```{r lists of outcomes for easy study}

# List of suicidal ideation variables
# Question Text: During the past 12 months, have you ever seriously thought about committing suicide?
suicidal_list <- c("h1su1","h3to130","h4se1","h5mn8")

# List of variables to indicate marijuana usage in the last 30 days
#During the past 30 days, on how many days did you use marijuana?
marijuana_outcome_use30_list <- c("h1to32","h3to110","h4to71","h5to21") 

# Total outcome variable list
outcome_vars_list <- c(suicidal_list, marijuana_outcome_use30_list)
```

## Biological Despair Variables 
Some predictors of biological despair need to be loaded separately from different files. Following are the functions to load them

### hba1c
```{r}
# This function takes in the full dataframe and path of the glucose biomarker file and adds the variables: hba1c
add_wave_4_glucose <- function(df, filebase = "~/gvfs/smb-share:server=vu1file.it.vanderbilt.edu,share=gaydosh_addhealth"){
  
  # Default file ending
  file_end = '/Gaydosh/Wave IV Biomarker Files/Wave IV Glucose/glu_a1c.xpt'
  
  #Create filepath
  filepath = str_c(filebase, file_end)
  
  glucose <- read_xpt(filepath)
  glucose <- glucose %>% janitor::clean_names() %>% select(aid, hba1c)
  df <- df %>% left_join(glucose, by = "aid") %>% select(aid, hba1c, everything())
  
  return(df)
}

# Example
# df <- df %>% add_wave_4_glucose() 
```

### crp
```{r}
# This function takes in the full dataframe and path of the crp_ebv biomarker file and adds the variable: crp
add_wave_4_crp <- function(df, filebase = "~/gvfs/smb-share:server=vu1file.it.vanderbilt.edu,share=gaydosh_addhealth" ){
  
  # Default file ending
  file_end = '/Gaydosh/Wave IV Biomarker Files/Wave IV CRP-EBV/crp_ebv.xpt'
  
  #Create filepath
  filepath = str_c(filebase, file_end)
  
  crp_df <- read_xpt(filepath)
  crp_df <- crp_df %>% janitor::clean_names() %>% select(aid, crp)
  df <- df %>% left_join(crp_df, by = "aid") %>% select(aid, crp, everything())
  
  return(df)
}

# Example
# df <- df %>% add_wave_4_crp()
```

### hdl, ldl, tg
```{r}
# This function takes in the full dataframe and path of the lipid biomarker file and adds the variables: hdl, ldl, tg
add_wave_4_lipids <- function(df, filebase = "~/gvfs/smb-share:server=vu1file.it.vanderbilt.edu,share=gaydosh_addhealth"){
  
  # Default file ending
  file_end = "/Gaydosh/Wave IV Biomarker Files/Wave IV Lipids/lipids.xpt"
  
  #Create filepath
  filepath = str_c(filebase, file_end)
  
  lipids <- read_xpt(filepath)
  lipids <- lipids %>% janitor::clean_names() %>% select(aid, hdl, ldl, tg)
  df <- df %>% left_join(lipids, by = "aid") %>% select(aid, hdl,ldl,tg, everything())
  
  return(df)
}

# Example
# df <- df %>%  add_lipids()
```

### wave 4 biological despair
```{r}
# This function takes in file paths for glucose, crp and lipid biomarker file paths for wave 4 and calls functions that loads these variables. 
# This abstracts the functionality of the individual functions that load the relevant variables
## Variables loaded with this function: hdl, ldl, tg, crp, hba1c
add_wave_4_bio_despair <- function(df, filebase = "~/gvfs/smb-share:server=vu1file.it.vanderbilt.edu,share=gaydosh_addhealth"){
  df <- df %>% 
    add_wave_4_glucose(filebase) %>% 
    add_wave_4_crp(filebase) %>% 
    add_wave_4_lipids(filebase)
  
  return(df)
}

# Example
# df <- df %>% add_wave_4_despair()
```


### h5bmi, h5waist
```{r}
# This function takes in the full dataframe and path of the anthro biomarker file and adds the variables: h5bmi, h5waist
add_wave_5_anthro <- function(df, filebase = "~/gvfs/smb-share:server=vu1file.it.vanderbilt.edu,share=gaydosh_addhealth"){
  
  # Default file ending
  file_end = "/Gaydosh/Wave V Biomarker Files/Wave V Anthropometrics/banthro5.xpt"
  
  #Create filepath
  filepath = str_c(filebase, file_end)
  
  anthro <- read_xpt(filepath) %>% 
    janitor::clean_names() %>% 
    select(aid, h5bmi, h5waist)
  
  df <- df %>% 
    left_join(anthro, by = "aid") %>% 
    select(aid, h5bmi, h5waist, everything())
  
}

# Example
# df <- df %>% add_wave_5_anthro()
```

### h5sbp, h5dbp, h5bpc1s4, h5bpjcls
```{r}
# This function takes in the full dataframe and path of the cardiovascular biomarker file and adds the variables: h5sbp, h5dbp, h5bpc1s4, h5bpjcls
add_wave_5_cardio <- function(df, filebase = "~/gvfs/smb-share:server=vu1file.it.vanderbilt.edu,share=gaydosh_addhealth" ){
  
  # Default file ending
  file_end = "/Gaydosh/Wave V Biomarker Files/Wave V Cardiovascular Measures/bcardio5.xpt"
  
  #Create filepath
  filepath = str_c(filebase, file_end)
  
  cardio <- read_xpt(filepath) %>% 
    janitor::clean_names() %>% 
    select(aid, h5sbp, h5dbp, h5bpcls4, h5bpjcls)
  
  df <- df %>% 
    left_join(cardio, by = "aid") %>% 
    select(aid, h5sbp, h5dbp, h5bpcls4, h5bpjcls, everything())
  
  return(df)
}

# Example
# df <- df %>% add_wave_5_cardio()
```

### Wave 5 biological despair
```{r}
# This function takes in file paths for antro and cardio vascular biomarker file paths and calls functions that loads these variables. 
# This abstracts the functionality of the individual functions that load the relevant variables
## Variables loaded with this function: h5sbp, h5dbp, h5bpc1s4, h5bpjcls, h5bmi, h5waist
add_wave_5_bio_despair <- function(df, filebase= "~/gvfs/smb-share:server=vu1file.it.vanderbilt.edu,share=gaydosh_addhealth") {
  
  df <- df %>% 
    add_wave_5_anthro(filebase) %>% 
    add_wave_5_cardio(filebase) 
  
  return(df)
}

# Example
# df <- df %>% add_wave_5_bio_despair()

```

### Add Biological Despair
```{r}
# Unified function to add all biological despair predictors to the dataset
add_bio_despair <- function(df, filebase= "~/gvfs/smb-share:server=vu1file.it.vanderbilt.edu,share=gaydosh_addhealth"){
  
  df <- df %>% 
    add_wave_3_bmi() %>%  # Function to create bmi in Wave 3 - this is present in 30-feature-engineering.Rmd file
    add_wave_4_bio_despair(filebase) %>%  # Add Wave 4 biological despair variables
    add_wave_5_bio_despair(filebase) # Add Wave 5 biological despair variables
  
  return(df)
}

```

## Lists of individual relevant variables
```{r parent}

parent_vars <- c("p_health", "pa10", "pa61", "pa62", "pa21", "pa55", "pa12", "pb8", "pa56", "pc31")

```

```{r wave 1}
w1_health <- c("h1to3", "h1to7", "h1gh1")

w1_social <- c("h1pf35", "h1pf36", "h1da7", "h1ed19", "h1ed20", "h1ed22", "h1fs13", "h1pr1", "h1pr2", "h1pr4", "h1nb3")

w1_peer <- c("h1su4", "h1su6", "h1su5", "h1su7")

w1_religion <- c("h1re3", "h1re4", "h1re7")

w1_family <- c("h1wp9", "h1wp10", "h1wp13", "h1wp14", "h1wp17f", "h1wp18f", "h1pf1", "h1pf4", "h1pf5", "h1pf23", "h1pf25", "h1pf24", "h1nm12f", "h1nm14", "h1nf12f", "h1nf14", "h1pr3", "h1pr5", "h1pr7", "h1pr8", "h1wp11", "h1wp12", "h1wp15", "h1wp16")

w1_parent_ses <- c("h1nf4", "h1nm4", "h1rf1", "h1rm1") 

w1_education <- c("h1gi20", "h1ee1", "h1ee2")

w1_prospects <- ("s45f")

w1_ses <- c(w1_parent_ses, w1_education, w1_prospects)

w1_ind_vars <- c(w1_health, w1_social, w1_peer, w1_religion, w1_family, w1_ses)
```

```{r wave 3}

w3_health <- c("h3to4", "h3to10", "h3gh18", "h3gh1", "h3hs1")

w3_social <- ("h3da15")

w3_prosocial <- c("h3cc6", "h3cc7", "h3cc8", "h3cc3")

w3_peer <- c("h3to133", "h3to135", "h3to136", "h3to134")

w3_religion <- c("h3re24", "h3re30", "h3re25", "h3re33", "h3re34")

w3_family <- c("close_mother", "close_father", "mother_support", "mother_support2", "father_support", "father_support2", "mother_warmth", "father_warmth", "h3wp65", "h3ma4", "h3ma3", "h3ma5", "h3ma6")

w3_structure <- c("h3cj160")

w3_partner <- c("h3mr1", "h3mr3_a", "h3mr6_a", "h3mr3_b", "h3mr6_b", "h3bm2")

w3_child <- c("h3mac10")

w3_education <- c("h3ed1", "h3ed19", "h3ed20", "h3ed27", "h3ed29", "h3ed30", "h3ed16", "h3ed17")

w3_poverty <- c("h3ec1c", "h3ec1d", "h3ec1e", "h3ec3", "h3ec4", "h3ec5", "h3ec6", "h3ec7", "h3ec9")

w3_employ <- c("industry_code3", "h3lm27")

w3_prospects <- c("h3ec59", "h3ec60")

w3_ses <- c(w3_education, w3_poverty, w3_employ, w3_prospects)

w3_ind_vars <- c(w3_health, w3_social, w3_prosocial, w3_peer, w3_religion, w3_family, w3_structure, w3_partner, w3_child, w3_ses)
```

```{r wave 4}
w4_health <- c("h4to3", "h4to6", "h4id7", "h4id9c", "h4gh8", "h4gh9", "h4gh1", "h4hs3")

w4_social <- c("h4ws4")

w4_prosocial <- c("h4da26", "h4da27")

w4_peer <- c("h4se4", "h4se5")

w4_religion <- c("h4re7", "h4re8", "h4re9", "h4re11")

w4_family <- c("h4wp24", "h4wp25", "h4wp38", "h4wp39", "h4wp23", "h4wp37", "h4ma1", "h4ma3", "h4ma5")

w4_structure <- c("h4wp3", "h4wp9", "h3wp16", "h3wp30")

w4_partner <- c("h4rd7b", "h4rd7d", "h4rd7e", "h4rd9", "h4rd12", "h4rd14")

w4_child <- c("h4tr9", "h4tr10")

w4_stress <-("h4mh29")

w4_education <- c("h4ed1", "h4ed2", "h4ed8", "h4ed9")

w4_poverty <- c("h4ec1", "h4ec3", "h4ec2")

w4_employ <- c("industry_code4", "h4lm23", "h4lm14", "h4lm20", "h4lm27")

w4_prospects <- c("h4ec19")

w4_ses <- c(w4_education, w4_poverty, w4_employ, w4_prospects)

w4_ind_vars <- c(w4_health, w4_social, w4_prosocial, w4_peer, w4_religion, w4_family, w4_structure, w4_partner, w4_child, w4_stress, w4_ses)
```

```{r wave 5}
w5_health <- c("h5to1", "h5to3", "h5id4", "h5da2", "h5id21", "h5id22", "h5id1", "h5lm13a", "h5id9")

w5_social <- c("h5ss6", "h5ss1", "h5ss2", "h5ss3a", "h5ss3b", "h5ss3c", "h5ss3d", "h5ss4a", "h5ss4b", "h5ss4c", "h5ss4d", "h5ss5a", "h5ss5b", "h5ss5c", "h5ss5d")

w5_prosocial <- c("h5ss8", "h5ss7")

w5_peer <- c("h5mn11", "h5mn12")

w5_religion <- c("h5re2", "h5re3")

w5_family <- c("h5wp14", "h5wp13", "h5wp28", "h5wp29", "h5wp27")

w5_structure <- c("h5wp15", "h5wp1", "h5wp4", "h5wp18")

w5_partner <- c("h5tr1", "h5tr6", "h5tr4", "h5hr1", "h5hr2", "h5tr3", "h5tr5", "h5tr7", "h5tr4_t", "h5tr5_t", "h5tr6_t", "h5tr7_t")

w5_child <- c("h5se21", "h5se22", "h5se23")

w5_stress <- c("h5mn5a", "h5mn5b", "h5mn5c", "h5mn5d", "h5mn5e", "h5mn6a", "h5mn6b", "h5mn6c", "h5mn6d", "h5mn6e", "h5mn6f","h5mn6g","h5mn6h","h5mn6i","h5mn6j","h5mn6k","h5mn6l","h5mn6m","h5mn6n")

w5_education <- c("h5od10", "h5od11")

w5_poverty <- c("h5ec1", "h5ec4", "h5ec2", "h5ec5a", "h5ec5b", "h5ec5c", "h5ec6")

w5_employ <- c("h5lm9", "h5lm10", "h5lm11", "h5lm12", "h5lm15", "h5lm5", "h5lm27")

w5_prospects <- c("h5ec9")

w5_ses <- c(w5_education, w5_poverty, w5_employ, w5_prospects)

w5_ind_vars <- c(w5_health, w5_social, w5_prosocial, w5_peer, w5_religion, w5_family, w5_structure, w5_partner, w5_child, w5_stress, w5_ses)
```


```{r vars by category, all waves}
ind_vars_health <- c(w1_health, w3_health, w4_health, w5_health)

ind_vars_social <- c(w1_social, w3_social, w4_social, w5_social)

ind_vars_prosocial <- c(w3_prosocial, w4_prosocial, w5_prosocial)

ind_vars_peer <- c(w1_peer, w3_peer, w4_peer, w5_peer)

ind_vars_religion <- c(w1_religion, w3_religion, w4_religion, w5_religion)

ind_vars_family <- c(w1_family, w3_family, w4_family, w5_family)

ind_vars_structure <- c(w3_structure, w4_structure, w5_structure)

ind_vars_partner <- c(w3_partner, w4_partner, w5_partner)

ind_vars_child <- c(w3_child , w4_child , w5_child)

ind_vars_stress <- c(w4_stress, w5_stress)

ind_vars_education <- c(w1_education, w3_education, w4_education, w5_education)

ind_vars_poverty <- c(w3_poverty, w4_poverty, w5_poverty)

ind_vars_employ <- c(w3_employ, w4_employ, w5_employ)

ind_vars_prospects <- c(w1_prospects, w3_prospects, w4_prospects, w5_prospects)

ind_vars_ses <- c(ind_vars_education, ind_vars_poverty, ind_vars_employ, ind_vars_prospects)

ind_vars_all <- c(parent_vars, w1_ind_vars, w3_ind_vars, w4_ind_vars, w5_ind_vars)

```