---
title: "10-import-data"
output:
  html_document: 
    code_folding: show
    theme: cosmo
    toc: yes
    toc_depth: 5
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r load libraries }
# Use pacman, which forces an install if the library isn't present on the running machine
if (!require("pacman"))
  install.packages("pacman")

# Load libraries required for this file
pacman::p_load(tidyverse, janitor, haven)
```

# Importing data
The purpose of this file is to contain all of the functions required or used to generate the dataset which will be investigated.  What follows are functions to import and generate data, and also pre-constructed lists for a simplified selection of features from the dataset.

## Sample workflow
The following code chunk defines the expected way that the following functions will be used, although other workflows are possible:

```{r, example chunk, eval=FALSE, purl=FALSE}
## load waves and join them
wave_data <- load_waves(1:5)
full_dataset <- get_working_dataset_full(wave_data, join_type = 'full')

## set outcome variable of interest
outcome = 'h5mn8'

## get the aids that you want
inner_aids <- get_inner(list(wave_data[[1]], wave_data[[3]], wave_data[[4]], wave_data[[5]]))

## use the features and ids that you want to select out what you want
working_ds <- full_dataset %>%
                filter(aid %in% inner_aids) %>%
                dplyr::select(aid, predictor_list, outcome)
```



## Public helper functions
These functions are helpers in case you need to know something about your dataframe more than what is provided in the general loading/joining interface.  Use as desired!

```{r generate unique list of vars from df}
# Return all unique values of var_name in data frames
get_unique_varnames <- function(df_list, var_name='aid'){
  # df_list : list of dataframes
  # var_name: variable of interest.  Expected in all lowercase letters
  # returns: list of distinct values of variable of interest across list of dataframes
  if(length(df_list)==0)
    stop("You didn't pass in any dataframes!")
  
  # Get list of var_names
  var_list <- df_list %>%
              bind_rows() %>%
              dplyr::select(var_name)
  
  # Return the unique ones
  return (distinct(var_list))
}
```

```{r check variable existence}
# Function to Check Existence of Variable in Data
# Use this function to check if a certain variable exists in a given. 
var_in_df <- function(df, varname){
  # df : dataframe to check in
  # varname : varaible to check for
  # return : Bool indicatind if the variable exists in the dataframe or not
  any(colnames(df) == varname)
}
## Examples of Usage : joined_waves %>% var_in("h4id5j")
```

```{r count responses by variable of interest}
# Function to check the counts by categories in the variables
count_response_by_variable <- function(df, ...){
  # df : dataframe where the variable exists
  # ... : variable you want to check the counts of - sent in as a string
  df %>%
    group_by_(...) %>% 
    summarise(n = n())
}

# Example of usage
# Tibble for first outcome with all of its predictors
# suicide_df <- joined_waves %>%
#   get_outcomes_df(outcome = "h5mn8", 
#                   predictors = setdiff(predictor_list, "aid"))
```

```{r returns intersection of variables in dataframes list}
# This function returns values that are common to all dataframes in list; 'aid' might be the only one here.
# Errors out if the variable isn't in all frames.

get_inner <- function(in_list, var_name='aid'){
  #in_list : contains lists of tibbles of wave data
  #var_name: column name of interest
  #returns: list of elements in the var_name column that are common to all of the dataframes
  all_var_list <- in_list %>%
    map(dplyr::select, var_name) %>%
    purrr::reduce(intersect) %>%
    pull(var_name)
  
  return(all_var_list)
}

## Example of usage
# wave_data <- load_waves(1:5)
# inner_aids <- get_inner(list(wave_data[[1]], wave_data[[3]], wave_data[[4]], wave_data[[5]]))
```

## Private helper functions
These helper functions are used within the publicly available functions.  Usage of these functions directly should be avoided.

## Public functions for loading data
These functions provide the overall functionality for loading a single file, multiple wave files, or multiple files in general using the dialog chooser.  Keep in mind that you can only use the dialog chooser on windows.  Use all functions as desired!

```{r load single file}
# This helper loads a single file given a full filepath f_path;

load_file <- function(f_path, use_clean_names = TRUE) {
  # f_path : filepath of AddHealth xpt file
  # returns: tibble of read data
  
  # Read the data
  read_data <- read_xpt(f_path)
  
  # Figure out what kind of data it is (currently race or a wave file)
  f_name = basename(f_path)
  
  # If it's wave data, extract the wave number and append it to the end of the dataframe.  Otherwise, read as normal.
  # We need to think through the consequences of this action.
  # if (str_extract(f_name, 'wave') == 'wave') {
  #   
  #   wave_no = parse_number(f_name)
  #   
  #   read_data <- read_data %>%
  #     mutate(wave_no = wave_no)
  # }
  
    # If you want clean names
  if(use_clean_names){
  read_data <- read_data %>%
              clean_names()
  }
  
  return(read_data)
}
```

```{r load using file chooser}
# This function allows choosing of files from a file chooser; returns a list containing the files.  Only available for Windows.

load_using_chooser <- function(use_clean_names = TRUE) {
  # use_clean_names: boolean of whether to fix column names
  # returns: list of tibbles of read data
  
  # Generate useful error for users not on Windows
  if(.Platform$OS.type != 'windows')
    stop(str_c("Sorry, this functionality isn't available on ", .Platform$OS.type))
  
  # Generate file chooser so wave or other data can be selected via GUI
  files <- choose.files(
    default = "",
    caption = "Select files",
    multi = TRUE,
    filters = Filters,
    index = nrow(Filters)
  )
  
  # Load data; returned as list of tibbles
  file_data <- files %>%
              map(load_file, use_clean_names)
  
  return(file_data)
}
```

```{r load all wave files easily}
# This function allows loading of files from a set location based on the waves of interest.

load_waves <- function(wave_no_list, use_clean_names = TRUE, filebase='~/gvfs/smb-share:server=vu1file.it.vanderbilt.edu,share=gaydosh_addhealth'){
  # wave_no_list : list of waves you want to load
  # use_clean_names: boolean of whether to fix (e.g., lowercase) column names
  # filebase : custom filepath for your local filesystem if needed; directly precedes .../Gaydosh/Core Files/...
  # returns: list of tibbles of wave files
  
  # Default file ending
  file_end = '/Gaydosh/Core Files/In Home Interview Files/wave'
  
  #Create filepath
  filepath = str_c(filebase, file_end)
  
  # Generate wave file list
  files <- str_c(filepath, wave_no_list, ".xpt")
  
  # Load data; returned as lists
  file_data <- files %>%
              map(load_file, use_clean_names)
  
  return(file_data)
}

# Example usage
# wave_data <- load_waves(1:5)
```

## Public functions for joining data and selection of data
These functions allow for joining data using different types of joins, and also allow for subsets of the data to be selected based on the input features
TODO:  We may need to think about what happens with the suffix - how will repeated variables (although we don't have that problem) be renamed?
TODO:  If you have names that are repeated, this will not work correctly because the variables will have repeated names (e.g., `h1fs1.x`) which won't match `h1fs1`.
TODO:  `get_working_dataset` requires assistance in determining how we can figure out which variables are factors without the codebook; note that somehow the `mutate_at` statement doesn't work in its current formulation (using | bars instead of commas) on ACCRE...

```{r perform join of data in an expedited way}
#This function returns the dataset with the selected predictors based on your join strategy.  However, with this function, you cannot easily retrieve the rows you left out without another join.

get_working_dataset_exp <- function(df_list, feature_list = NULL, join_type = 'inner', by='aid'){
  # df_list : list of dataframes
  # feature_list : list of desired features.  If null, select all features.
  # join_type : type of join you'd like to perform
  # by: key on which the join should be performed
  # returns: full joined tibble with selected features
  
  # Obtain correct join type based on inputs
  if(join_type=='inner')
    jtype <- inner_join
  else if(join_type=='left')
    jtype <- left_join
  else if(join_type=='right')
    jytpe <- right_join
  else if(join_type=='full')
    jtype <- full_join
  else{
    stop(str_c('Join type', by, 'is not supported.', sep=' '))
  }
  
  # Do the join
  ds <- reduce(df_list, jtype, by=by)
  
  # Remedy null feature list if null
  if (is.null(feature_list)){
    feature_list = colnames(ds)
  }
  
  # Make sure all values in feature list are in the column names
  missing_features <- setdiff(feature_list, colnames(ds))
  
  if(length(missing_features) != 0){
    stop('The following features are missing in the set of dataframes you passed in: ', str_c(missing_features, collapse=', '))
  }
  
  # Fix factor issues and dplyr::select features
  ds <- ds %>% 
    mutate_at(vars(starts_with("h"), starts_with("r"), starts_with("m")), as.factor) %>%
    dplyr::select(feature_list)
  
  return(ds)
}

```

This function has the same problems as `get_working_dataset_exp`.  However, this function may end up being more useful depending on the file format that we decide to save the fully joined data in (if we can).  The reading of the remote file is still TODO.
TODO: Something is strange with my mutate_at function; it can only work with commas and not bars.

```{r perform join of data}
# This functions generates the data by starting with an outright full join of the data

get_working_dataset_full <- function(df_list, feature_list = NULL, join_type = 'inner', load_join=FALSE){
  # df_list : list of dataframes
  # feature_list : list of desired features.  If null, select all features.
  # join_type : type of join you'd like to perform
  # by: UNIMPLEMENTED, TODO
  # load_join: if true, the fully joined data will be loaded (UNIMPLEMENTED, TODO)
  # returns: full joined tibble with selected features
  
  # Do the full join
  if(load_join){
    # Here, we want to load the joined data from file if possible and contractually acceptable.
  }
  else
    ds <- reduce(df_list, full_join, by='aid')
  
  # Extract AIDs from each df
  aid_lists <- df_list %>% map(dplyr::select, aid)
  
  # Obtain correct join type based on inputs
  if(join_type=='inner')
    aids <- reduce(aid_lists, intersect) %>% pull(aid)
  else if(join_type=='left')
    aids <- aid_lists[[1]] %>% pull(aid)
  else if(join_type=='right')
    aids <- aid_lists[[length(aid_lists)]] %>% pull(aid)
  else if(join_type=='full')
    aids <- reduce(aid_lists, union) %>% pull(aid)
  else{
    stop(str_c('Join type', by, 'is not supported.', sep=' '))
  }
  
  # Filter out aids
  ds <- ds %>%
    filter(aid %in% aids)
  
  # Remedy null feature list if null
  if (is.null(feature_list)){
    feature_list = colnames(ds)
  }
  
  # Make sure all values in feature list are in the column names
  missing_features <- setdiff(feature_list, colnames(ds))
  
  if(length(missing_features) != 0){
    stop('The following features are missing in the set of dataframes you passed in: ', str_c(missing_features, collapse=', '))
  }
  
  # Fix factor issues and select features
  ds <- ds %>% 
    mutate_at(vars(starts_with("h"), starts_with("r"), starts_with("m")), as.factor) %>%
    dplyr::select(feature_list)
  
  return(ds)
}

# Example usage:
# wave_data = load_data(1:5)
# full_dataset <- get_working_dataset_full(wave_data, join_type = 'full')
```

TODO: We need to make some decisions about how demographic data will likely be included in the generated data.  Where are these variables included in the variable lists and how can we most intuitively integrate them into our data loading?
```{r}
# This function will be used to add age and race demographic into the data
add_demographics <- function(df){
  # df : joined waves input
  # return : new data frame with race and age information at each wave. 
  #          age_w{x} - where x is the wave number (e.g. age_w5, age_w4..etc)
  # race info consists of three variables - mulrace1, race1, race1rul
  df %>% 
    mutate(birth_in_months = (as.double(paste("19",h1gi1y,sep=""))*12 )+ as.double(h1gi1m)) %>% 
    mutate(age_w1 = floor((((as.double(paste("19",iyear,sep=""))*12 )+imonth) - birth_in_months)/12) ) %>% 
    mutate(age_w3 = floor((((iyear3*12)+imonth3)-birth_in_months)/12)) %>% 
    mutate(age_w4 = floor((((iyear4*12)+imonth4)-birth_in_months)/12)) %>% 
    mutate(age_w5 = floor((((iyear5*12)+imonth5)-birth_in_months)/12)) %>% 
    left_join(race, by ="aid") %>% 
    dplyr::select(-one_of(c('birth_in_months'))) 
}
```



## Binarize "times used" outcome variables by 0 vs rest of values
Some outcome variables are not binary. This function is used to binarize outcome variables by 0 versus rest of values.
```{r}
binarize_var <- function(df, variable){
  
  df[[variable]] <- as.character(df[[variable]])
  df[[variable]] <- as.numeric(df[[variable]])
  df[[variable]] <- ifelse(df[[variable]] > 0, 1, 0)
  df[[variable]] <- as.character(df[[variable]])
  df[[variable]] <- as.factor(df[[variable]])
  
  return(df[[variable]])
}



# Example:
# test_data<-depression
# test_data$h1fs1
# test_data[["h1fs1"]]
# test_data[["h1fs1"]] <- binarize_var(test_data,"h1fs1")

```





## Lists of Selected Predictor Variables
The following are some lists of predictor variables, organized by area of interest.  Note that the AIDs have been removed from all of these.

```{r lists of predictors for easy study}
# List of all Anxiety Variables - Note only wave 4 is being looked at for Anxiety measures
anxiety_list <- c("h4pe6","h4pe14","h4pe22","h4pe30")
diagnosed_anxiety_list <- c("h4id5j","h5id6i")


# List of all Optimism Variables
optimism_list <- c("h4pe7","h5pe1","h4pe15","h5pe2","h4pe23","h4pe31","h5pe3")


# List of all Depression Variables
depression_list <- c(
                     "h1fs1","h1fs3","h1fs4","h1fs5","h1fs6","h1fs7","h1fs11","h1fs15","h1fs16","h1fs17",   # Wave 1 ~ 10 vars
                     "h3sp5","h3sp6","h3sp7","h3sp8","h3sp9","h3sp10","h3sp11","h3sp12","h3sp13",           # Wave 3 ~ 9 vars
                     "h4mh18","h4mh19",'h4mh20','h4mh21',"h4mh22","h4mh23","h4mh24","h4mh26","h4mh27",      # Wave 4 ~ 9 vars
                     "h5ss0a","h5ss0b","h5ss0c","h4mh25","h5ss0d")                                          # Wave 5 ~ 5 vars

diagnosed_depression_list <- c("h4id5h","h5id6g")

# List of all Demographic Variable
demographic_list <- c("race1","mulrace1","race1rul","bio_sex")

demographic_list_comp <- c("age_w1", "age_w3", "age_w4", "age_w5")

# Total predictor variable list
predictor_list <- c(anxiety_list, diagnosed_anxiety_list, optimism_list, depression_list, diagnosed_depression_list)

# Note: Look into importing these later
#c4var009,
#c4var010,
#c4var002
```

# Lists of Selected Outcome Variables
The following are some lists of outcome variables, organized by area of interest.  Note that the AIDs have been removed from all of these.
```{r lists of outcomes for easy study}

# List of suicidal ideation variables
# Question Text: During the past 12 months, have you ever seriously thought about committing suicide?
suicidal_list <- c("h1su1","h3to130","h4se1","h5mn8")

# List of variables to indicate marijuana usage in the last 30 days
#During the past 30 days, on how many days did you use marijuana?
marijuana_outcome_use30_list <- c("h1to32","h3to110","h4to71","h5to21") 

# Total outcome variable list
outcome_vars_list <- c(suicidal_list, marijuana_outcome_use30_list)
```
