---
title: "21-explore-1"
output: html_notebook
---

# Optimism Varible Questions

## Summary of Optimism Subset

```{r}
summary(optimism)
```

```{r}
describe(optimism)
```


## Missing Value Visualization

```{r}
optimism %>% 
  vis_miss()
```

```{r}
optimism %>% 
  gg_miss_var(show_pct = T)
```



## Histogram of  Variables

```{r}
hist.data.frame(optimism)
```

### Optimism Q: I'm always optimistic about my future
```{r}
optimism %>%
  select(h4pe7,h5pe1) %>% 
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()
```

### Optimism Q: I hardly ever expect things to go my way

```{r}
optimism %>%
  select(h4pe15,h5pe2) %>% 
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()
```

### Optimism Q: I expect more good things to happen to me than bad

```{r}
optimism %>%
  select(h4pe23,h5pe3) %>% 
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()
```

### Optimism Q: I rarely count on good things happening to me

Only exists in Wave 5
```{r}
optimism %>%
  select(h4pe31) %>% 
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()

```

# Visualization of Wave 4 Variables and Survey Answers for those

```{r}
optimism_sub <- optimism %>% 
  dplyr::select(h4pe7,h4pe15,h4pe23,h4pe31)


optimism_sub <- data.frame(lapply(optimism_sub, factor, ordered=TRUE, 
                          levels=1:7, 
                          labels=c("1","2",
                                   "3","4","5","6","8")))

likert(optimism_sub)

```

```{r}
plot(likert(optimism_sub, ))
```

```{r}
plot(likert(optimism_sub),
     type="heat",
           low.color = "white",
           high.color = "blue",
           text.color = "black",
           text.size = 4,
           wrap = 50)
```


```{r}
plot(likert(optimism_sub),
     type="density",
           facet = TRUE,
           bw = 0.5)
```


