---
title: "40-h2o-helpers"
output: html_document
editor_options: 
  chunk_output_type: console
---

This notebook details functionality regarding h2o starting and ending. 

```{r load libraries}
library(pacman)
pacman::p_load(h2o, stringr)
```

The following function will generate the bash command that we need to use to start an instance of h2o.
```{r function to make command string}
make_cmd <- function(port_assign){
  java_dir <- '/accre/arch/easybuild/software/BinDist/Java/11.0.2/bin/java'
  h2o_dir <- '~/R/x86_64-pc-linux-gnu-library/3.6/h2o/java/h2o.jar'
  #cmd_str <- str_c(java_dir, '-ea -cp -jar ', h2o_dir, ' water.H2OApp -hash_login -login_conf realm.properties -port ', port_assign)
  cmd_args <- c('-Xmx32g', '-ea', '-cp', h2o_dir, "water.H2OApp", '-hash_login', '-login_conf', 'realm.properties', '-port', port_assign)
  
  return(list(cmd <- java_dir,
              args <- cmd_args))
}
```


The following function randomly assigns a port number and tries to start up h2o on that port.
```{r make command from random port no}

try_h2o <- function(log_out, log_err){
  
  #generate some random number for the port start
  port_no <- sample(40000:60000, 1)
  
  #generate the command to run to start R
  cmd <- make_cmd(port_no)
  
  #run the command and don't wait for it to finish
  print(str_c('Trying to start h2o on port ', port_no, ' ....this will take at least 15 seconds'))
  system2(cmd$cmd, args=cmd$args, stdout=log_out, stderr=log_err, wait=FALSE)
  Sys.sleep(15)
  
  return(port_no)
}
  
```

The following function creates temp files and is copied from h2o's h2o-3/connection.R implemention.
```{r}
get_tempfile <- function(temp_type){
  
  #Make sure the file type is correct
  if(missing(temp_type) || !(temp_type %in% c("stdout", "stderr", "pid")))
    stop("type must be one of 'stdout', 'stderr', or 'pid'")
  
  #Generate the directory name based on user credentials
  if(.Platform$OS.type == "windows") {
    usr <- gsub("[^A-Za-z0-9]", "_", Sys.getenv("USERNAME", unset="UnknownUser"))
  } else {
    usr <- gsub("[^A-Za-z0-9]", "_", Sys.getenv("USER", unset="UnknownUser"))
  }

  #Generate the log file name based on user credentials
  temp_dir_path <- tempfile()
  dir.create(temp_dir_path)
  if(temp_type == "stdout")
    file.path(temp_dir_path, paste("h2o", usr, "started_from_r.out", sep="_"))
  else if(temp_type == "stderr")
    file.path(temp_dir_path, paste("h2o", usr, "started_from_r.err", sep="_"))
  else
    file.path(temp_dir_path, paste("h2o", usr, "started_from_r.pid", sep="_"))
}
```


If the port is in use, this will make our h2o server generation fail.  So, if it does fail, then we need to try to connect on a different port.  After the h2o instance is successfully connected using h2o.init(), then we have the port number to use for all future instances.

```{r securely start h2o}
start_h2o <- function(){
  
  #make sure relevant files exist, otherwise, we need to not move forward.
  if(!(file.exists('realm.properties') & file.exists('.h2oconfig'))){
    error("Looks like you haven't created your credentials for running h2o.  Please generate these credentials now!")
  }
  
  #Create temp files
  err_file <- get_tempfile("stderr")
  out_file <- get_tempfile("stdout")
  
  #signals
  success <- FALSE
  tries <- 10
  
  #try to start a server of h2o
  port_no <- try_h2o(out_file, err_file)
  
  #keep trying to connect until h2o_init() succeeds or we run out of tries
  while((success==FALSE) & (tries>0)){
    
    tryCatch({
      #try to start h2o; this will fail if the port was never created
      h2o.init(port=port_no)
      success=TRUE
    }, error = function(e){
      
      #if we did fail, let's try another port.
      port_no <- try_h2o(out_file, err_file)
      tries <- tries - 1
      
      #print something
      print(str_c('Failed to connect on port ', port_no, '.  Trying ', tries, ' more times...'))
    })
  }
  
  #print status after we're done
  if (success==TRUE){
    print('H2o successfully started!')
  } else {
    stop('H2o could not start!')
  }
  
  return(port_no)
  
} 
```

