---
title: "41-modeling-RF"
output: html_notebook
---

The purpose of this document is to build a randomforest that models the outcomes: suicidal ideation, marijuana usage and alcohol abuse. 

#Steps Taken

1. Splitting the data. [x]
2. Bootstrapping the training set n times (n = 1000). [x]
3. Splitting the bootstrap set into a training a validation set. [x]
4. Creating n different models. [x]
- finding the best tuning parameters for each model using grid search (Tuning parameters: ntrees and max_depth).[x]
- creating a final model (using k-fold cross validation) that uses the best tuning parameters from the previously conducted grid search.[x]
- fitting the model onto the validation set created earlier.[x]
5. Saving all models and the performance metrics into a list. [x]
6. Evaluation of all model parameters and the stability of those parameters.  



```{r load libraries}
library(data.table)
library(mlr)
library(splitstackshape) #stratified function
library(boot)
library(tidyverse)
library(rsample)
```


```{r random forest model function}
# Note: The steps are not in order because the function needed to be created first.
rf_model <- function(#training_frame,         
  #validation_frame，
  train,          # training set or bootstrap
  k = 5,          # ks of k fold
  max_depths_tune = c(100,150,200), # vector of max_depths to test on
  ntree_tune = c(5,10,25), # vector of ntrees to test on
  outcome)
  {     
  
  
    # create lists to save models and results
    model_list <- list() 
    results_list <- list()
    
  
    # Step 3
    # Using splitstackshape::stratified to create a stratified split on the outcome variable - 90-10 split
    straps_splits <- stratified(as.data.frame(train), outcome, 0.9, bothSets = T) 
    
    # pull out training and validation frames
    training_frame <- as.h2o(straps_splits[[1]])
    validation_frame <- as.h2o(straps_splits[[2]])
    
    # Step 4
    # Create all combinations of tuning parameters
    params <- expand.grid(max_depth = max_depths_tune, ntrees = ntree_tune)
    model_tune_list <- rep(NA, nrow(params))
    
    #initialize function parameters for h2o.randomForest
    outputs <- "h5mn8"
    inputs <- setdiff(predictor_list, "aid")
    
    #Training models on all combinations of parameters
    for(i in 1:nrow(params)){
      rf_i <- h2o.randomForest(
        training_frame = training_frame,
        x= inputs,
        y= outputs, 
        model_id = "rf_covType_v1",
        max_depth = params[i,][[1]],
        ntrees = params[i,][[2]], 
        stopping_rounds = 2,
        nfolds = 5,
        fold_assignment = "Stratified",
        score_each_iteration = T,
        seed = 42
      )
      
      #Test on validation frame
      result <- h2o.performance(rf_i, as.h2o(validation_frame))
      
      #Pulling model metric (AUC)
      model_tune_list[i] <- result@metrics$AUC
    }
    
    # Pull the index of the tuning parameter which produced the highest AUC
    best_index <- which(model_tune_list==max(model_tune_list))[1]
    
    #Train final model based on best tuning parameters
    rf_i <- h2o.randomForest(
      training_frame = training_frame,
      x=inputs,
      y=outputs, 
      model_id = "rf_covType_v1",
      ntrees = params[best_index,2], 
      max_depth = params[best_index,1],
      stopping_rounds = 2,
      nfolds = k,
      fold_assignment = "Stratified",
      score_each_iteration = T,
      seed = 42
    )
    
    
    return(list(rf_i, h2o.performance(rf_i, validation_frame)))
}

```




```{r}
# # Creating the subset of the dataset we will be using for the model
# #Note:
# #1. Have not added age, race and sex/gender information in this subset as of yet. 
# #2. Initially creating the model to predict suicidal ideation at wave 5. 
# #3. Initially making the decision to remove all NA values in wave 5.
# # Step 1
# data_subset_list <- setdiff(predictor_list, "aid") %>% append("h5mn8") # list of variables to use 
# data_subsets <- joined_waves[data_subset_list] %>% na.omit() # subsetting the data based on the variables
# splits <- stratified(data_subsets, c("h5mn8"), 0.7, bothSets = T) 
# train <- splits[[1]]
# test <- splits[[2]]
# 
# 
# # Step 2
# n = 2 # input 
# straps <- train %>% bootstraps(times = n, strata = "h5mn8")
# 
# combined_models <- vector(mode = "list", length = n)
# combined_performances <- vector(mode = "list", length = n)
# 
# 
# # These are the tuning parameters that we decide to tune.
# max_depths <- c(5,10,25)
# ntrees <- c(100,150,200)
# 
# 
# for(i in 1:length(straps$splits)){
#   results <- as.data.frame(straps$splits[[i]])%>%
#     rf_suicide()
#   
#   print(results)
#   
#   combined_models[i] <- results[1]
#   combined_performances[i] <- results[2]
#   
# }
# combined_models[[1]]
# 
# typeof(data_subsets)
# 
# first <- rep(NA,n)
# second <- rep(NA,n)
# third <- rep(NA,n)
# fourth <- rep(NA,n)
# fifth <- rep(NA,n)
# sixth <- rep(NA,n)
# seventh <- rep(NA,n)
# eighth <- rep(NA,n)
# ninth <- rep(NA,n)
# tenth <- rep(NA,n)
# 
# 
# 
# for(i in 1:n){
#   first[i] <- data.frame(h2o.varimp(combined_models[[i]]))[1,1]
#   second[i] <- data.frame(h2o.varimp(combined_models[[i]]))[2,1]
#   third[i] <- data.frame(h2o.varimp(combined_models[[i]]))[3,1]
#   fourth[i] <- data.frame(h2o.varimp(combined_models[[i]]))[4,1]
#   fifth[i] <- data.frame(h2o.varimp(combined_models[[i]]))[5,1]
#   sixth[i] <- data.frame(h2o.varimp(combined_models[[i]]))[6,1]
#   seventh[i] <- data.frame(h2o.varimp(combined_models[[i]]))[7,1]
#   eighth[i] <- data.frame(h2o.varimp(combined_models[[i]]))[8,1]
#   ninth[i] <- data.frame(h2o.varimp(combined_models[[i]]))[9,1]
#   tenth[i] <- data.frame(h2o.varimp(combined_models[[i]]))[10,1]
# } 
# 
# as.data.frame(table(first))
# as.data.frame(table(second)) %>% arrange(-Freq)
# as.data.frame(table(third)) %>% arrange(-Freq)
# as.data.frame(table(fourth)) %>% arrange(-Freq)
# as.data.frame(table(fifth)) %>% arrange(-Freq)
# as.data.frame(table(seventh)) %>% arrange(-Freq)
# as.data.frame(table(eighth)) %>% arrange(-Freq)
# as.data.frame(table(ninth)) %>% arrange(-Freq)
# as.data.frame(table(tenth)) %>% arrange(-Freq)
# 
# ## First Position
# 
# as.data.frame(table(first)) %>% 
#   ggplot(aes(x = first, y = Freq))+
#   geom_col() +
#   coord_flip()
# 
# ## Second Position
# 
# as.data.frame(table(second)) %>% 
#   arrange(-Freq) %>% 
#   ggplot(aes(x = reorder(second, Freq), y = Freq))+
#   geom_col() +
#   coord_flip()
# 
# 
# ## Third Position
# 
# as.data.frame(table(third)) %>% 
#   arrange(-Freq) %>% 
#   ggplot(aes(x = reorder(third, Freq), y = Freq))+
#   geom_col() +
#   coord_flip()
# 
# ## Fourth Position
# 
# as.data.frame(table(fourth)) %>% 
#   arrange(-Freq) %>% 
#   ggplot(aes(x = reorder(fourth, Freq), y = Freq))+
#   geom_col() +
#   coord_flip()
# 
# 
# ## Fifth Position
# 
# 
# as.data.frame(table(fifth)) %>% 
#   arrange(-Freq) %>% 
#   ggplot(aes(x = reorder(fifth, Freq), y = Freq))+
#   geom_col() +
#   coord_flip()
# 
# 
# 
# 
# data.frame(table(as.matrix(rbind(first, second, third, fourth, fifth, sixth, seventh, eighth, ninth, tenth)))) %>% 
#   ggplot(aes(x = reorder(Var1, Freq), y = Freq))+
#   geom_col() +
#   coord_flip()
# 
# 
# 
# 
# 
# 
# ## Final Model
# #This is created based on the top features extracted from the previous section.
# 
# final_features <- c("h5ss0b", "h5ss0a", "h5pe1", "h5id6g", "h5ss0d", "h5pe3", "h5pe2", "h5ss0c", "h4pe7", "h1fs17", "h5id6i", "h4pe14", "h4mh27", "h4pe23","h4pe22", "h1fs11", "h4pe31", "h4pe30", "h4mh24", "h3sp8")
# ## These features will be extracted from the previous section - hard coded for now. 
# 
# tuning_param <- expand.grid(max_depth = c(7,8,9,10), ntrees = c(100,150,200,250))
# model_tune_list <- rep(NA, nrow(tuning_param))
# 
# for(i in 1:nrow(tuning_param)){
#   rf_i <- h2o.randomForest(
#     training_frame = as.h2o(train),
#     x=final_features,
#     y="h5mn8", 
#     model_id = "rf_covType_v1",
#     max_depth = tuning_param[i,][[1]],
#     ntrees = tuning_param[i,][[2]], 
#     stopping_rounds = 2,
#     nfolds = 5,
#     fold_assignment = "Stratified",
#     score_each_iteration = T,
#     seed = 42
#   )
#   
#   result <- h2o.performance(rf_i, as.h2o(train))
#   
#   model_tune_list[i] <- result@metrics$AUC
# }
# 
# best_index <- which(model_tune_list==max(model_tune_list))[1]
# 
# 
# rf_final_model <- h2o.randomForest(
#   training_frame = as.h2o(train),
#   x=final_features,
#   y="h5mn8", 
#   model_id = "rf_covType_v1",
#   ntrees = tuning_param[best_index,][[2]], 
#   max_depth = tuning_param[best_index,][[1]],
#   stopping_rounds = 2,
#   nfolds = 5,
#   fold_assignment = "Stratified",
#   score_each_iteration = T,
#   seed = 42
# )
# rf_final_model
# 
# h2o.varimp(rf_final_model)
# 
# perf <- h2o.performance(rf_final_model, as.h2o(test))
# perf
# 
# plot(perf,type = "roc")
# 
# h2o.confusionMatrix(perf)
```


