---
title: "60-results"
output: html_document
---

The purpose of this document is to explore the results of the models built in the 50 series.

```{r load libraries}
library(pacman)
pacman::p_load(tidyverse, h2o, yardstick)
```

In general, one shouldn't use these functions directly.  They're helpers to perform the functionality of the main methods.
```{r helper functions}

```

The following define the main functions which will be used in the 60 series for further investigation and the 70 series to understand the results of the analysis.

```{r main functions}
# This function returns the metrics for a set of models on an input dataframe
get_metric_set <- function(df, model_list){
  # df: Training or testing frame
  # model_list: list of h2o model objects
  # return: tibble with all applicable metrics for each model tested on df

  # get metric set
  metric_set <- model_list %>%
  map_df(get_performance, test_data = df) %>% 
  mutate(model = factor(1:length(model_list))) %>% 
  select(model, everything())
   
  # return metric set
  return(metric_set)
}
```



