---
title: "73-experiments-prob-drinking"
output:
  html_notebook:
    code_folding: hide
    theme: lumen
    toc: yes
    toc_float: yes
    toc_depth: 4
  html_document:
    df_print: paged
    toc: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r import source files, warning=FALSE, message=FALSE}
source("function_import.R")
```

# Exploring Problematic Drinking

**Purpose.** In this work, we will explore the relation between identified measures of despair of interest (e.g., personality measures of self-consciousness, individual and composite item scores from the CES-D assessment) and descriptors of diseases of despair.  We will achieve this goal through modeling the outcomes based on the included predictors, and robustly assess the importance of the included features in predicting the outcomes via bootstrapping.  We will use two well-known machine learning models, random forests and LASSO, which are both frequently used to measure the relative importance of the predictors included in the models.  Lastly, we'll generate trained and tuned models using this reduced feature set which can be used by others wish to predict the identified outcomes.

**Subject inclusion.** For this investigation, we will omit the entirety of Wave 2.  This is commonly done in analyses of AddHealth data due the design of the original study.  Otherwise, our dataset will include only subjects who have predictor and outcome data in _all_ of the waves.

**Outcome variables.** We will model _problematic drinking_ assessed at Wave 5.  

**Predictor variables.** The predictors for these models are hand-picked, and based on previous work, relevance, and subject matter expertise. The set of predictors and the set of outcomes are disjoint.  Predictors from Waves 1-5 (excluding Wave 2, see above) are included, and will be detailed in the following analysis.

```{r load libraries, include=FALSE}
# Use pacman, which forces an install if the library isn't present on the running machine
if (!require("pacman")) install.packages("pacman")
p_install(plotly, force = FALSE)
pacman::p_load(tidyverse, h2o, furrr)
#pacman::p_load(tidyverse, janitor, rsample, h2o, furrr, rcompanion, ggcorrplot)

```

```{r initializations, include=FALSE}
h2o.init() 
h2o.no_progress()
future::plan(multiprocess)
```

```{r seeds for reproducibility}
seed= 3895
set.seed(seed)
```

# Dataset generation

The predictors we will be using will be the the variable `predictor_list` loaded from `10-import-data.Rmd` file. These initial set of predictors will be based of the list of variables that describe anxiety, depression, and optimism.

```{r load raw data and formulate dataset, warning=TRUE, message=TRUE}
## set outcome variable of interest
outcome = 'h5to15'
filebase = '/scratch/p_gaydosh_lab'

#create data in specified form
dataset_list <- generate_datasets(outcome, binarize=TRUE, filebase=filebase, seed_val=seed)

#parse out dataset components
wave_data <- dataset_list$wave_data
full_dataset <- dataset_list$full_dataset
ds_raw <- dataset_list$ds_raw_outcome
ds <- dataset_list$ds_final

#ml splits of the data
training_df <- dataset_list$training_df
validation_df <- dataset_list$validation_df
testing_df <- dataset_list$testing_df

```


## Binarizing and recoding details

This is the table with all the counts for different levels of the variables. We can see 97 is a legitimate skip variable. We can send this into the binarize_var function and it will be coded as 0 (not responding to questions involving problematic drinking).

```{r warning=FALSE, message=FALSE}
ds_raw %>% 
  group_by(.data[[outcome]]) %>% 
  summarise(total = n(), type = class(.data[[outcome]]))

ds_raw %>% explore_outcome(outcome, binary = TRUE)
```


```{r warning=FALSE, message=FALSE}
ds_raw %>% 
  binarize_var(outcome, legit_skip = T, skip_var = 97) %>% 
  group_by(.data[[outcome]]) %>% 
  summarise(total = n(), type = class(.data[[outcome]]))

ds_raw %>%
  explore_outcome(outcome, binary = FALSE, legit_skip = TRUE, skip_var = 97)

```
This shows the result of recoding; 0 and legitimate skip (97) are recoded as the negative class, and the rest are regarded as the positive class.  The results are consistent with the previous cell; 1372 97s + 3802 0s = 5174 total negative class.  This fully accounts for the 9,349 subjects in the current dataframe.

```{r warning=FALSE, message=FALSE}
ds %>% explore_outcome(outcome)
```
Now that the dataframe has been binarized and the null 11 outcomes dropped, There are 9338 subjects in the dataset with 76 total predictors, aid, and the outcome variable.  The distribution of the data is a slight imbalance, but not by much.  I think this warrants using AUC as the optimization and selection metric.

# Data exploration and visualization
Here, we comment about the general characteristics of the data based on the provided visualizations.  We comment on missingness of data, any strange or unusual behavior (e.g., strong imbalances), and any correlation that sticks out.

```{r warning=FALSE, message=FALSE}
#Report about the characteristics of the subjects left out of the join
ds %>% explore_dropped()

```

```{r eda distributions, fig.width = 15, fig.height=15}
# Visualize distributions of variables of interest
ds %>% 
  dplyr::select(-aid) %>%
  graph_bar_discrete(df = .,
                     plot_title = "Distributions of Discrete Variables",
                     max_categories = 50,
                     num_rows = 3,
                     num_cols = 3,
                     x_axis_size = 12,
                     y_axis_size = 12,
                     title_size = 15)
```

```{r eda missing}
ds %>%
  graph_missing(only_missing = TRUE,
                title = "Percent Missing",
                box_line_size = .5,
                label_size = .5,
                x_axis_size = 12,
                y_axis_size = 12,
                title_size = 15)

```

```{r eda correlations}
ds %>%
  #dplyr::select(1:20) %>%
  pairwise_cramers_v() %>%
  plot_cramer_v(x_axis_angle = 90,
                plot_title = "Association among Categorical Variables",
                interactive = TRUE)
```
We can see that some of the categorical variables actually have non-trivial correlation with others.  It appears as if entire blocks of predictors have high correlation among themselves.  (h4id5j, h4id5h) have correlation of 0.6, (h3id5j, h4mh26) have correlation of 0.51, and correlations among the h4mh* and h1fs* variables are near 0.5 or higher.  This may result in importance sharing in the future, and we may need to decide to remove some of these variables.

# Robust feature evaluation {.tabset .tabset-fade .tabset-pills}
In this section, we explore the importance of features using bootstrapping.

## RF model
The RF models are chosen based on a grid search using the following the parameters: 

- max depth: maximum depth allowed for a single tree in the RF  
- number of trees: maximum number of trees allowed in the RF  
- mtries: the number of columns sampled for each tree split  
- stopping_rounds: number of rounds to continue when stopping_metric doesn't improve, based on a simple moving average
- min_rows: the minimum number of rows required to split the internal node
- balance classes: whether to balance the classes or not
- stopping_metric: metric which results in early stopping of training of the model
- categorical encoding: use one hot encoding to create straightforward comparison with LASSO

Note that here, if you don't set stopping rounds, early stopping will not be enabled.

```{r bootstrap rf training, warning=FALSE, message=FALSE}
rf_params <- list(max_depth = 50,
                  ntrees = 150,
                  mtries = c(-1, 20, by=5),
                  min_rows = c(2,10, by=2),
                  stopping_rounds = 4,
                  balance_classes = FALSE,
                  stopping_metric = 'auc',
                  categorical_encoding = 'one_hot_explicit')


# define number of bootstraps
n_boot = 50

boots_rf <- model_feature_selection("RF", training_frame = training_df,
                                        validation_frame = validation_df,
                                        hyper_params = rf_params,
                                        selection_metric = 'AUC',
                                        outcome = outcome, n = n_boot, seed=seed)

```


The following table displays the mean performance metrics for the bootstrapped models on the validation set, removing values for which there are NA.

```{r evaluate bootstrap model performance rf, warning=FALSE, message=FALSE}
mean_bs_rf_perf <- get_metric_set_from_perfs(boots_rf$perfs) %>%
  dplyr::select(accuracy, mpce, sens, spec, ppv, npv, roc_auc, pr_auc,
                tns, tps, fns, fps, no_n, no_p,  err_rate, bal_accuracy, everything()) %>%
  summarise_if(is.numeric, mean, na.rm=TRUE) %>%
  mutate(model = 'bs_rf') %>%
  dplyr::select(model, everything())

mean_bs_rf_perf
```


```{r}
boot_rf_mdi <- boots_rf$mdi %>%
  get_median_placement(use_base_var = TRUE) %>%
  add_attribute_names('predictor', full_dataset) %>%
  dplyr::select(predictor, att_name, overall_rank)

head(boot_rf_mdi, 20)
```

This table returns variable importance ranks that returned from each of the bootstrapped models.

```{r fig.width = 10, fig.height = 12}
# Needs to be fixed so that axes don't overlap each other and obscure understanding
plot_placement_boxplot(boots_rf$mdi)
```

Now, let's look at the permutation importance:
```{r}
boot_rf_perm_plt <- boots_rf$models %>%
  get_aggregated_permute_imp(training_df, outcome=outcome)
```

```{r}
met <- 'roc_auc'
boot_rf_perm <- boot_rf_perm_plt %>%
  get_permute_placement(metric_oi=met) %>%
  add_attribute_names('predictor', full_dataset) %>%
  dplyr::select(predictor, everything())

head(boot_rf_perm, 20)
```

Now, let's look at these models together:
```{r fig.width = 16, fig.height = 14}
cbind(boot_rf_mdi[1:20,], dplyr::select(boot_rf_perm[1:20,], -met))
```

```{r fig.width = 10, fig.height = 12}
plot_permute_var_imp(boot_rf_perm, metric = roc_auc)
```

## LASSO model
In this step, we model the relation between the outcomes and the predictors using a linear regression with L2 regularization.  This drives the importance of unimportant and redudant features towards zero.

```{r feature selection lasso, include=FALSE}
# Function parameters
lasso_params <- list(alpha = c(1))
# Call modeling function using function parameters and show visualization of results.  Recommend the number of features that should be used.  Report performance metric stats.

boots_lasso <- model_feature_selection( "Lasso",
                                            training_frame = training_df,
                                            validation_frame = validation_df,
                                            hyper_params = lasso_params,
                                            outcome = outcome, 
                                            n = n_boot)
```


```{r evaluate bootstrap model performance lasso, warning=FALSE, message=FALSE}
mean_bs_lasso_perf <- get_metric_set_from_perfs(boots_lasso$perfs) %>%
  dplyr::select(accuracy, mpce, sens, spec, ppv, npv, roc_auc, pr_auc,
                tns, tps, fns, fps, no_n, no_p,  err_rate, bal_accuracy, everything()) %>%
  summarise_if(is.numeric, mean, na.rm=TRUE) %>% 
  mutate(model='bs_lasso') %>%
  dplyr::select(model, everything())

mean_bs_lasso_perf
```
### Feature importances: LASSO
#### Coefficient-based variable importance

```{r}
boot_lasso_mdi <- boots_lasso$mdi %>%
  get_median_placement(use_base_var = TRUE) %>%
  add_attribute_names('predictor', full_dataset) %>%
  dplyr::select(predictor, att_name, overall_rank)

head(boot_lasso_mdi, 20)
```

```{r fig.width = 10, fig.height = 12}
plot_placement_boxplot(boots_lasso$mdi)
```

#### Permutation importance
```{r}
boot_lasso_perm_plt <- boots_lasso$models %>%
  get_aggregated_permute_imp(training_df, outcome=outcome)
```

```{r aggregate lasso permutations and get metrics}
boot_lasso_perm <- boot_lasso_perm_plt %>%
  get_permute_placement(metric_oi=met) %>% #set in random forest section
  add_attribute_names('predictor', full_dataset) %>%
  dplyr::select(predictor, everything())

head(boot_lasso_perm, 20)
```

```{r fig.width = 10, fig.height = 12}
plot_permute_var_imp(boot_lasso_perm, metric = roc_auc)
```


#### Coefficient vs. Permutation importance
Now, we compare the feature importances generated by the two different approaches.  The traditional method of evaluating feature importance for regression methods is through analysis of the coefficients.

```{r fig.width = 16, fig.height = 14}
cbind(boot_lasso_mdi[1:20,], select(boot_lasso_perm[1:20,], -met))
```

## Comparison: Model Type Mean Performance
The following table compares the mean performance of bootstrapped random forests to the mean performance of bootstrapped LASSO methods.
```{r}
bs_comp_perfs <- rbind(mean_bs_rf_perf, mean_bs_lasso_perf) 
bs_comp_perfs
```



## Comparison: Model Type Feature Importance
Here, we look at the aggregated results of the bootstrapped predictors and compare the models generated to each other.
```{r}
joined_results <- boot_rf_perm %>%
  dplyr::select(-met) %>%
  full_join(dplyr::select(boot_lasso_perm, -met), by=c("predictor", "att_name"), suffix=c('.rf', '.lasso')) %>%
  mutate(mean_rank = rowMeans(dplyr::select(., overall_rank.rf, overall_rank.lasso), na.rm=TRUE)) %>%
  arrange(mean_rank)

head(joined_results, 20)
```
The following visualization provides the intuition about the differences in the rankings between model types.  They're ordered by the overall mean importance, and for a given variable, the differences in rank are shown.
```{r fig.width = 20, fig.height = 30}
# Comparison of top_n features
joined_results %>%
  compare_feature_select(interactive = TRUE,
                         top_n = 100,
                         opacity = 0.50,
                         plot_title = "Permutation Importance of Predictors by Model")
```

# Generation of final model {.tabset .tabset-fade .tabset-pills}

## RF model
In this step, we build the final model for the random forest.  We use slightly more values in order to come up with the best model, keeping in mind the number of combinations that are required to run to evaluate the grid.

```{r final model evaluation rf}

# # Spans of hyper parameters for random forest
rf_params <- list(max_depth = 50,
                  ntrees = 150,
                  mtries = seq(-1, 30, by=5),
                  min_rows = seq(5, 60, by=5),
                  balance_classes = c(TRUE, FALSE),
                  stopping_metric = 'AUCPR',
                  categorical_encoding = 'one_hot_explicit')

# rf_params <- list(max_depth = seq(20, 50, 20),
#                   balance_classes = TRUE,
#                   categorical_encoding= 'one_hot_explicit')

# Function parameters
final_model_rf <- rf_model(outcome,
                           training_frame = training_df,
                           validation_frame = validation_df,
                           nfolds = 5,
                           hyper_params = rf_params, model_seed=seed)

```
### Performance
The final random forest performance metrics are shown below:

```{r}
# show model final performance
print(final_model_rf[[2]])
```

### Features: permutation importance

```{r get plottable permutation importance, message=FALSE, warning=FALSE}
final_rf_perm_plt <- c(final_model_rf[[1]]) %>%
  get_aggregated_permute_imp(training_df, outcome=outcome)
```


```{r get metrics for final rf}
final_rf_perm <- final_rf_perm_plt %>%
  get_permute_placement(metric_oi=met) %>%
  add_attribute_names('predictor', full_dataset) %>%
  dplyr::select(predictor, everything())

head(final_rf_perm, 20)
```

```{r plot rf final permutation, fig.width = 10, fig.height = 12}
plot_permute_var_imp(final_rf_perm, metric = roc_auc)
```
### Comparison with bootstrap results
This section investigates the differences in the bootstrap results vs the features generated from the random forest final model.  The following table shows the overall differences in rank.

```{r}
rf_joined_results <- final_rf_perm %>%
  dplyr::select(-met) %>%
  full_join(dplyr::select(boot_rf_perm, -met), by=c("predictor", "att_name"), suffix=c('.final', '.bootstrap')) %>%
  mutate(mean_rank = (overall_rank.final + overall_rank.bootstrap)/2) %>%
  arrange(mean_rank)

head(rf_joined_results, 20)
```

The following plot provides visualizations for the difference in the final model rankings vs the bootstrap.

```{r fig.height=14, fig.width=16}
# Comparison of top_n features
rf_joined_results %>%
  compare_feature_select(sel_cols = c("overall_rank.final", "overall_rank.bootstrap"),
                         interactive = TRUE,
                         top_n = 100,
                         opacity = 0.50,
                         plot_title = "Permutation Importance of Predictors: Final vs. Bootstrap")
```


## LASSO model
Now, we create the final model for LASSO.  There is no substantial difference between this method and the bootstrap methods, other than the data upon which the model is being built.
```{r final model evaluation lasso}
# Function parameters
lasso_params <- list(alpha = c(1))

final_model_lasso <- lasso_model(training_frame = training_df,
                                 validation_frame = validation_df,
                                 outcome = outcome,
                                 nfolds = 5,
                                 hyper_params = lasso_params)
```

The final LASSO performance metrics are shown below:
```{r}
# show model final performance
print(final_model_lasso[[2]])
```

### Features: permutation importance
```{r get plottable permutation importance lasso, message=FALSE, warning=FALSE}
final_lasso_perm_plt <- c(final_model_lasso[[1]]) %>%
  get_aggregated_permute_imp(training_df, outcome=outcome)
```

```{r get permute features for final lasso}
final_lasso_perm <- final_lasso_perm_plt %>%
  get_permute_placement(metric_oi=met) %>%
  add_attribute_names('predictor', full_dataset) %>%
  dplyr::select(predictor, everything())

head(final_lasso_perm, 20)
```
```{r plot lasso final permutation, fig.width = 10, fig.height = 12}
plot_permute_var_imp(final_lasso_perm, metric = roc_auc)
```


### Comparison with bootstrap results
This section investigates the differences in the bootstrap results vs the features generated from the LASSO final model.  The following table shows the overall differences in rank.

```{r}
lasso_joined_results <- final_lasso_perm %>%
  dplyr::select(-met) %>%
  full_join(dplyr::select(boot_lasso_perm, -met), by=c("predictor", "att_name"), suffix=c('.final', '.bootstrap')) %>%
  mutate(mean_rank = (overall_rank.final + overall_rank.bootstrap)/2) %>%
  arrange(mean_rank)

head(lasso_joined_results, 20)
```

The following plot provides visualizations for the difference in the final model rankings vs the bootstrap.

```{r fig.width = 20, fig.height = 20}
# Comparison of top_n features
lasso_joined_results %>%
  compare_feature_select(sel_cols = c("overall_rank.final", "overall_rank.bootstrap"),
                         interactive = TRUE,
                         top_n = 100,
                         opacity = 0.50,
                         plot_title = "Permutation Importance of Predictors: Final vs. Bootstrap")
```


## Comparison: Final model features
Here, we compare the features generated by the permutation importance between the two final models.

```{r}
rf_lasso_final_joined_results <- final_rf_perm %>%
  dplyr::select(-met) %>%
  full_join(dplyr::select(final_lasso_perm, -met), by=c("predictor", "att_name"), suffix=c('.rf', '.lasso')) %>%
  mutate(mean_rank = (overall_rank.rf+overall_rank.lasso)/2) %>%
  arrange(mean_rank)

head(rf_lasso_final_joined_results, 20)
```

The following visualization provides the intuition about the differences in the rankings between the final model types.  They're ordered by the overall mean importance, and for a given variable, the differences in rank are shown.

```{r fig.width = 16, fig.height = 14}
# Comparison of top_n features
rf_lasso_final_joined_results %>%
  compare_feature_select(sel_cols = c("overall_rank.rf", "overall_rank.lasso"),
                         interactive = TRUE,
                         top_n = 100,
                         opacity = 0.50,
                         plot_title = "Permutation Importance of Predictors: Random Forest vs Lasso")
```

## Comparison: Final model performance
With the final models generated, we're now able to compare their performance metrics.
```{r final model comparison, fig.width = 10, fig.height = 12}
# Comparison of performance metrics
valid_perf <- get_metric_set_from_perfs(perf_list = list(final_model_rf[[2]], final_model_lasso[[2]])) %>%
  mutate(model = c('rf', 'lasso'))

testing_perf <- get_metric_set_from_models(testing_df, list(final_model_rf[[1]], final_model_lasso[[1]]), out=outcome) %>%
  mutate(model = c('rf', 'lasso'))
```

**Validation and selection.**
The following table shows the comparison between models in terms of the validation set.  We can select our final model based on the best performing model according to the metric.
```{r}
print(valid_perf)
```

**Testing performance.**
The following shows the performance of both the models on the test set.  Note that although we don't use this test set to evaluate the final models, we can still see how our selected method would have performed.
```{r}
print(testing_perf)
```

The following plots show a comparison between the performance of the models on the validation and test sets.  Again, we don't choose the model based on the test set, but curiosity dictates that we view this performance.

```{r fig.width = 10, fig.height = 6}
# Show plots side by side
metrics_of_interest = c('model', 'accuracy', 'bal_accuracy', 'mpce', 'sens', 'spec', 'ppv', 'npv', 'pr_auc', 'roc_auc')
valid_plt <- plot_metric_set(dplyr::select(valid_perf, all_of(metrics_of_interest)), plot_title = "Model comparison for validation set")
test_plt <- plot_metric_set(dplyr::select(testing_perf, all_of(metrics_of_interest)), plot_title = "Model comparison for testing set")
gridExtra::grid.arrange(gridExtra::arrangeGrob(valid_plt, test_plt, ncol=2, nrow=1))
```


