---
title: "utils"
output: html_document
---

#Utility functions
The purpose of this notebook is to provide utility functions that assist in the analysis, but are also completely tangential to the analysis.

```{r}
library(pacman)
pacman::p_load(arrow, tidyverse)
```
We'll be using arrow as a file format here due to its fast and robust performance, as well as its fidelity in retaining datatypes from the initial dataframe.  For more information, see [here](https://arrow.apache.org/docs/r/articles/arrow.html).

## Test filepath saving location and parquet
```{r}

save_filepath_remote <- 'G:/DSI/test.parquet'
save_filepath_local <- './test.parquet'
test_iris <- iris %>% tibble()

#Can we write CSVs?
write_csv(test_iris, path=save_filepath_local)
#write_csv(test_iris, path=save_filepath_remote)

#Can we write parquets?
write_parquet(test_iris, path=save_filepath_local)
#write_parquet(test_iris, path=save_filepath_remote)

```
Can write locally, but not remotely.  Probably an issue with file permissions.

## Create fully joined dataset and save to vu1file
```{r create-and-save-full-join}
source('function_import.R');
```

```{r}
wave_data <- load_waves(1:5, filebase='G:/')
full_dataset <- get_working_dataset_full(wave_data, join_type = 'full')
```

```{r}
save_filepath <- 'G:/DSI/wave1-5_full'
```

